### 2022-04-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@b82956c6e0ed0ab5ef5a4d80a336662ab4fb5fb4 🚀

\hypertarget{host-distribution}{%
\subsection{Host distribution}\label{host-distribution}}

Bats are found worldwide and are one of the most diverse groups among
mammals (Moratelli \& Calisher, 2015), and one of the main animal
reservoir for different strains of betacoronaviruses (Drexler et al.
2014). This has attracted attention to areas where high diversity of
bats, and therefore presumably high diversity of betacoronaviruses, can
be an important issue for human health (Calisher et al. 2006, Moratelli
and Calisher 2015). Accordingly, we collected the IUCN rangemaps for
known hosts of betacoronaviruses, to illustrate where hotspots of host
diversity are. These results are presented in Fig xx.a. As per our
current knowledge of which bats are hosts of betacoronaviruses, these
hotspots are primarily South-East Asia, parts of Europe, and to a lesser
extent sub-saharan Africa. Even the subset of chiroptera that are hosts
of betacoronaviruses fits the evolutionary timeline of the group.
Chiropterans can be classified as Microchiroptera and macrochiroptera,
where macrochiroptera have an older history from an evolutionary
perspective compared to macrochiroptera (Teeling et al. 2005, Springer
2013).

@@

South-East Asia has a high diversity of bats (\textbf{Kingston, 2010}),
and our results show that part of that diversity includes
betacoronavirus hosts. High density of hosts sharing the same virus
(albeit possibly different strains) calls into question the evolution of
the bat antiviral immune system and its co-evolution with viruses, which
may result in distinct immunological responses in different areas, as
evidenced in other bat species (Banerjee et al. 2020). Immune
characteristics that allow bats to be better adapted to infection by
emerging viruses (Gorbunova et al., 2020; Irving et al., 2021) may be
related to a wide variety of diets (Jones et al., 2022; Moreno Santillán
et al., 2021; Banerjee et al., 2020; Schneeberger et al., 2013).
Considering whether viruses easily adapted to multiple hosts have lower
virulence on these hosts, or lower ability to jump to hosts with
different immune characteristics, should yield valuable additional
predictors for the total risk of spillover. Previous research
(\textbf{Anthony et al., 2017; Mollentze \& Streicker, 2020}) states
that locally diverse bat communities could maintain more viruses and
hence, a higher probability of having a pathogen that could represent a
risk for human health; locally diverse, virus-rich bats communities
could represent an increased risk of spillover under climate change
(\textbf{Ice ice berg berg}). This probability involves multiple
factors, among which the relatedness of hosts (which can make the jumps
easier (\textbf{Longdon et al., 2011; Mollentze et al., 2020; Wolfe et
al., 2007}), and the overall tendency of hosts within a locality to
share viruses, which may limit viral diversity because of within-host
competition (Leeks et al., 2018; Sallinen et al., 2020). All things
considered, the richness of known betacoronaviruses hosts is not a
sufficient predictor of spillover risk.

\hypertarget{viral-evolutionary-distinctiveness}{%
\subsection{Viral evolutionary
distinctiveness}\label{viral-evolutionary-distinctiveness}}

Higher host diversity may not result in a higher viral diversity, for
example if all hosts share the same viruses, or share closely
evolutionarily related strains. For this reason, we quantified and
mapped the evolutionary distinctiveness of betacoronaviruses, based on
their position in a molecular phylogeny. Viral evolutionary
distinctiveness largely tracks host diversity, particularly in southern
China but, oddly, not throughout the rest of southeast Asia. This
indicates, perhaps, that many distinctive viruses remain to be
discovered in this region (an idea that is unsurprising given the
growing realization, around the emergence of SARS-CoV-2, that a unique
lineage of similar viruses are widespread in bats but still mostly
undescribed). The most distinct betacoronaviruses are found in South
America, a region with a comparatively lower number of hosts; this
suggests that the South American bat-betacoronvirus complex has been
more isolated, and is probably undergoing a different co-evolutionary
dynamic. Alternatively, this distinctiveness hostpot may be a product of
under-sampling: South-America is one of the places where the fewest
betacoronaviruses have been discovered (Anthony et al. 2017), and adding
more viruses would bring the distinctiveness of known sequences down.
Previous work has suggested the Americas may be a hotspot of both
undiscovered bat viruses in general (Olival et al. 2017, Allen et al.
2017) and coronavirus specifically (Anthony et al. 2017), though not
necessarily betacoronaviruses, and particularly not those in clades with
notable zoonotic potential.

@@

\hypertarget{geographic-mosaic-of-bat-betacoronavirus-risk}{%
\subsection{Geographic Mosaic of bat-betacoronavirus
risk}\label{geographic-mosaic-of-bat-betacoronavirus-risk}}

@@

\hypertarget{global-distribution-of-spillover-risk}{%
\subsection{Global distribution of spillover
risk}\label{global-distribution-of-spillover-risk}}

>>>>

\hypertarget{host-distribution-does-not-match-virus-distinctiveness}{%
\subsection{Host distribution does not match virus
distinctiveness}\label{host-distribution-does-not-match-virus-distinctiveness}}

Bats are found worldwide and are one of the most diverse groups among
mammals (Moratelli \& Calisher, 2015), and one of the main animal
reservoir for different strains of betacoronaviruses (Drexler et al.
2014). This has attracted attention to areas where high diversity of
bats, and therefore presumably high diversity of betacoronaviruses, can
be an important issue for human health (Calisher et al. 2006, Moratelli
and Calisher 2015). Accordingly, we collected the IUCN rangemaps for
known hosts of betacoronaviruses, to illustrate where hotspots of host
diversity are. These results are presented in fig.~\ref{fig:richness}
(top). As per our current knowledge of which bats are hosts of
betacoronaviruses, these hotspots are primarily South-East Asia, parts
of Europe, and to a lesser extent parts of Africa in the -25-0 range of
latitudes. Previous research (\textbf{Anthony et al., 2017; Mollentze \&
Streicker, 2020}) states that locally diverse bat communities could
maintain more viruses and hence, a higher probability of having a
pathogen that could represent a risk for human health; locally diverse,
virus-rich bats communities could represent an increased risk of
spillover under climate change (\textbf{Ice ice berg berg}). This
probability involves multiple factors, among which the relatedness of
hosts (which can make the jumps easier (\textbf{Longdon et al., 2011;
Mollentze et al., 2020; Wolfe et al., 2007}), and the overall tendency
of hosts within a locality to share viruses, which may limit viral
diversity because of within-host competition (Leeks et al., 2018;
Sallinen et al., 2020). All things considered, the richness of known
betacoronaviruses hosts is not a sufficient predictor of spillover risk,
and we should move towards a more coevolutionary understanding of this
system.

@@

High density of hosts sharing the same virus (albeit possibly different
strains) calls into question the evolution of the bat antiviral immune
system and its co-evolution with viruses, which may result in distinct
immunological responses in different areas, as evidenced in other bat
species (Banerjee et al. 2020). Immune characteristics that allow bats
to be better adapted to infection by emerging viruses (Gorbunova et al.,
2020; Irving et al., 2021) may be related to a wide variety of diets
(Jones et al., 2022; Moreno Santillán et al., 2021; Banerjee et al.,
2020; Schneeberger et al., 2013). Considering whether viruses easily
adapted to multiple hosts have lower virulence on these hosts, or lower
ability to jump to hosts with different immune characteristics, should
yield valuable additional predictors for the total risk of spillover.
The subset of chiroptera that are hosts of betacoronaviruses fits the
evolutionary timeline of the group. Chiropterans can be classified as
Microchiroptera and macrochiroptera, where macrochiroptera have an older
history from an evolutionary perspective compared to macrochiroptera
(Teeling et al. 2005, Springer 2013). Specifically, we would expect that
the New World group of bats, being more evolutionary distinct, would
also have evolutionary distinct viruses. Indeed fig.~\ref{fig:richness}
(bottom) reveals it to be the case. By contrast, South-Eastern Asia has
a lot of non-evolutionary distinct bats, but evolutionary-distinct
viruses.

Viral evolutionary distinctiveness largely tracks host diversity,
particularly in southern China but, oddly, not throughout the rest of
southeast Asia. This indicates, perhaps, that many distinctive viruses
remain to be discovered in this region (an idea that is unsurprising
given the growing realization, around the emergence of SARS-CoV-2, that
a unique lineage of similar viruses are widespread in bats but still
mostly undescribed). The most distinct betacoronaviruses are found in
South America, a region with a comparatively lower number of hosts; this
suggests that the South American bat-betacoronvirus complex has been
more isolated, and is probably undergoing a different co-evolutionary
dynamic. Alternatively, this distinctiveness hostpot may be a product of
under-sampling: South-America is one of the places where the fewest
betacoronaviruses have been discovered (Anthony et al. 2017), and adding
more viruses would bring the distinctiveness of known sequences down.
Previous work has suggested the Americas may be a hotspot of both
undiscovered bat viruses in general (Olival et al. 2017, Allen et al.
2017) and coronavirus specifically (Anthony et al. 2017), though not
necessarily betacoronaviruses, and particularly not those in clades with
notable zoonotic potential.

\hypertarget{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}{%
\subsection{The phylogeographic regions of hosts and their viruses
overlap}\label{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}}

@@

\hypertarget{coevolution-informed-spillover-risk-is-different-in-space}{%
\subsection{Coevolution-informed spillover risk is different in
space}\label{coevolution-informed-spillover-risk-is-different-in-space}}

@@

\hypertarget{human-occupancy-drives-different-levels-of-effective-risk-globally}{%
\subsection{Human occupancy drives different levels of effective risk
globally}\label{human-occupancy-drives-different-levels-of-effective-risk-globally}}
